---
title: "phoenix_ucr_pdf_processing"
output: pdf_document
date: "2024-04-16"
---
```{r}
#install.packages("pdftools")
library(pdftools)
library(tidyverse)
library(stringr)
```

pdf(file = "tmp.pdf")
plot(1, main = "mytext")
dev.off()
pdftools::pdf_text(pdf = "tmp.pdf")

```{r}
phx_2022 <- pdf_text("2022_AZ_Phoenix_UCRIndexCount.pdf")
phx_2023 <- pdf_text("2023_AZ_Phoenix_UCRIndexCount.pdf")
```
check the column names
```{r}
head(phx_2022)
#head(phx_2023)
```

how many elements did this create?
```{r}
length(phx_2022)
#length(phx_2023)

```

```{r}
phx_ucr22 <- phx_2022[1]
phx_ucr22 <- strsplit(phx_ucr22, "\n")
phx_ucr22 <- phx_ucr22[[1]]

#phx_ucr23 <- phx_2023[1]
#phx_ucr23 <- strsplit(phx_ucr23, "\n")
#phx_ucr23 <- phx_ucr23[[1]]

```

what does this look like?
```{r}
head(phx_ucr22)
```
ditch the white space
```{r}
phx_ws <- trimws(phx_ucr22)

```
let's get rid of the extra text
```{r}
phx_cleaner <- phx_ws[grep("Criminal Homicide", phx_ws):
                                   grep("Total Property Crimes", phx_ws)]
```

check it out
```{r}
phx_cleaner_still <- str_split_fixed(phx_cleaner, " {2,}",18)
phx_cleaner_still
```
make it a table
```{r}
phx_table <- data.frame(phx_cleaner_still)
#crime type; jan; feb; mar; q1; apr; may; jun; q2; jul; aug; sep; q3; oct; nov; dec; q4; ytd
```


larceny parsed funny - let's investigate
```{r}
larceny_elements <- phx_cleaner[19]
larceny_elements <- str_squish(larceny_elements)
larceny_clean <- str_split_fixed(larceny_elements, " {1,}",20)
larceny_clean
```



add some column names
```{r}
names(phx_table) <- c("crime_type",
                                                  "jan",
                                                  "feb",
                                                  "mar", 
                                                  "q1",
                                                  "apr",
                                                  "may",
                                                  "jun",
                                                  "q2",
                                                  "jul", 
                                                  "aug",
                                                  "sep",
                                                  "q3",
                                                  "oct",
                                                  "nov",
                                                  "dec", 
                                                  "q4",
                                                  "ytd"
                      )
```
let's see it now
```{r}
phx_table
```
get rid of old rows & unnecessary columns
```{r}

phx_table_subset <- phx_table[c(1,4,7,12,15,19,20,24),c(1:4,6:8,10:12, 14:16) ]
```
turn the crime type variable into a factor and recode
```{r}
#convert to factor first
phx_table_subset$crime_type <- as.character(phx_table_subset$crime_type)

#try again...
phx_table_subset$crime_type <- recode(phx_table_subset$crime_type, 
                                "1. CRIMINAL HOMICIDE" = "HOMICIDE",
                                "2. RAPE TOTAL" = "RAPE",
                                "3. ROBBERY TOTAL"= "ROBBERY",
                                "4. ASSAULT TOTAL"= "ASSAULT",
                                "BURGLARY" = "BURGLARY",
                                "6. LARCENY - THEFT TOTAL" = "LARCENY",
                                "7. MOTOR VEHICLE THEFT TOTAL" = "MVT"
                                                                                  )
```
write it out!
```{r}
write.csv(phx_table_subset, "phx_ucr22.csv")
#write.csv(phx_table_subset, "phx_ucr23.csv")

```




#notes
#Online version
#pdftools::pdf_text(pdf = "http://arxiv.org/pdf/1403.2805.pdf")
