---
title: "rhode island ucr processing"
output: pdf_document
date: "2024-05-01"
---
```{r}
library(lubridate)
library(tidyverse)
library(stringr)
```
these are monthly files so let's read them all in using a list command
```{r}
# Set the path to the directory containing the CSV files
folder_path <- "C:\\OneDrive\\OneDrive - ahdatalytics.com\\Clients\\Real Time Crime Index\\Open Source Data\\Rhode Island\\State UCR"

# List all CSV files in the directory
file_list <- list.files(path = folder_path, pattern = "\\.csv$", full.names = TRUE)

# Function to read and process each file
process_file <- function(file_name) {
  # Read the CSV file
  df <- read_csv(file_name, show_col_types = FALSE)
  
  # Assuming the date column is named 'Date'
  # Convert to date format if not already and ensure the date format matches your data
  df$`Incident Date` <- as.Date(df$`Incident Date`, format = "%m/%d/%Y")  # Adjust format as needed
  
  # Create month and year columns
  df$Month <- month(df$`Incident Date`, label = TRUE, abbr = TRUE)  # Returns abbreviated month name
  df$Year <- year(df$`Incident Date`)
  
  return(df)
}

# Apply the function to each file and store results in a list of data frames
ldf <- lapply(file_list, process_file)
```

what does the top look like
```{r}
head(ldf[1])
```
create a new table that aggregates by agency + month + year + UCR.Offense.Code
```{r}
# Define the function to aggregate counts
aggregate_counts <- function(ldf) {
  # Ensure the data is grouped and summarised correctly
  aggregated_data <- ldf %>%
    group_by(`Agency Name`, `UCR Offense Code`, Month, Year) %>%
    summarise(count = n(), .groups = 'drop')  # Drop grouping for easier use later
  
  # Return the aggregated data
  return(aggregated_data)
}

# Assuming ldf is a list of data frames
monthly_counts <- lapply(ldf, aggregate_counts)

```
check the new list
```{r}
head(monthly_counts[1])
```


let's get the crimes as columns
```{r}
#create a function that will transform all df's in the list to wide
transform_to_wide <- function(df) {
  # Pivot the dataframe to a wider format
  df_wide <- df %>%
    pivot_wider(names_from = `UCR Offense Code`, values_from = count)

  return(df_wide)
}

#apply function to list
ldf_wide <- lapply(monthly_counts, transform_to_wide)
# Print the resulting dataframe
print(ldf_wide)
```
remove records with NA in the year column - convert NA counts to 0 - also converts month field to numeric representation
jan ==1, feb ==2, etc.
```{r}
#function to do both
clean_df <- function(df) {
  # Filter out rows where 'year' and 'month' column has NA values
  clean_data <- df %>%
    filter(!is.na(Year) | !is.na(Month))
  
  # Convert all columns (except the first one) to numeric and replace NA with 0
  clean_data1 <- clean_data %>%
    mutate(across(.cols = -1, ~ as.numeric(replace(., is.na(.), 0))))
  
  return(clean_data1)
}

#apply function to dataframes in list
ldf3 <- lapply(ldf_wide, clean_df)
print(ldf3)

```
sum column values to get SRS figures from NIBRS
From NIBRS OFFENSE CODES MANUAL 2011
```{r}
#murder 09A - Murder & Nonnegligent Manslaughter

#rape - 11A - Forcible Rape

#agg assault - 13A - Aggravated Assault

#robbery - 120 - Robbery

#burglary - 220 - Burglary/Breaking & Entering

#Larceny/Theft Offenses to sum
#Pocket-picking Property,  Purse-snatching Property,  Shoplifting Property,  Theft From Building Property,  Theft From Coin-Operated Machine or Device Property,  Theft From Motor Vehicle Property,  Theft of Motor Vehicle Parts or Accessories Property,All Other Larceny Property

process_larceny <- function(df) {
  # Define the columns related to larceny
  columns_to_larc <- c(
    "Pocket-picking", "Purse-snatching", "Shoplifting", "Theft From Building",
    "Theft From Coin-Operated Machine or Device", "Theft From Motor Vehicle",
    "Theft From Motor Vehicle Parts/Accessories", "All Other Larceny"
  )
  
  # Check if all specified columns exist in the dataframe
  missing_cols <- setdiff(columns_to_larc, names(df))
  if (length(missing_cols) > 0) {
    warning("The following columns are missing in the dataframe and will be excluded from the sum: ", paste(missing_cols, collapse=", "))
    columns_to_larc <- setdiff(columns_to_larc, missing_cols)
  }
  
  # Calculate the sum of the specified columns if any exist
  if (length(columns_to_larc) > 0) {
    df$Larceny <- rowSums(df[, columns_to_larc], na.rm = TRUE)
  } else {
    df$Larceny <- 0  # Set Larceny to 0 if no columns are available for summing
  }
  
  return(df)
}

#mvt - 240 - Motor Vehicle Theft

#arson -200 - Arson

ldf3_larc <- lapply(ldf3, process_larceny)
print(ldf3_larc)
```
drop unnecessary columns
```{r}
 # Columns to keep
  columns_to_keep <- c("Agency Name", "Murder and Nonnegligent Manslaughter",
                       "Rape", "Robbery", "Aggravated Assault", "Burglary/Breaking and Entering",
                       "Larceny", "Motor Vehicle Theft", "Arson", "Month", "Year")

  # New column names
  final_col_names <- c("Agency Name","Murder", "Rape", "Robbery",
                       "Agg Assault", "Burglary", "Larceny", "MVT", "Arson", "Month", "Year")

# Define the function to clean and rename dataframe columns
clean_and_rename <- function(df) {

  # Subset the dataframe
  df <- df[, columns_to_keep, drop = FALSE]  # Ensure it returns a dataframe even if only one column

  # Rename the columns
  names(df) <- final_col_names

  return(df)
}
```
rename each column for all dataframes in list
```{r}
# Apply the processing function to each dataframe in the list
ldf4 <- lapply(ldf3_larc, clean_and_rename)

# Example to print the head of the first cleaned and renamed dataframe
head(ldf4[[1]])
```


row bind the list together
```{r}
combined_df <- do.call(rbind, ldf4)

```
write it out!
```{r}
write.csv(combined_df, "or_state_Jan22_Present.csv")


```




#notes
#Online version
#pdftools::pdf_text(pdf = "http://arxiv.org/pdf/1403.2805.pdf")
