---
title: "rtci_sample_audit_vizready"
output: github_document
date: 07-09-2024
---
load up libraries
```{r}
library(lubridate)
library(tidyverse)
library(stringr)
library(readr)
library(dplyr)
```
let's load in the sample from the collector script
```{r}
samp <- read_csv("C:\\OneDrive\\OneDrive - ahdatalytics.com\\Clients\\Real Time Crime Index\\Open Source Data\\Collected Sample Data\\rtci_sample_extracted.csv")    
```
function to standardize the month column
```{r}
# Process the month and year columns
samp <- samp %>%
  mutate(
    Month = case_when(
      is.na(Month) ~ NA_real_,
      !is.na(as.numeric(Month)) ~ as.numeric(Month),
      !is.na(match(tolower(Month), tolower(month.abb))) ~ match(tolower(Month), tolower(month.abb)),
      !is.na(match(tolower(Month), tolower(month.name))) ~ match(tolower(Month), tolower(month.name)),
      TRUE ~ NA_real_
    ),
    Year = as.numeric(Year)
  )

# Dataframe for invalid months (NA values in Month column)
invalid_months <- samp %>%
  filter(is.na(Month) | Month < 1 | Month > 12)

# Dataframe for invalid years - we use UCR for 2022
invalid_years <- samp %>%
  filter(is.na(Year) | Year < 2023 | Year > 2025)

# Display the dataframes
print("Invalid Months:")
print(invalid_months)

print("Invalid Years:")
print(invalid_years)

# Just keep the valid month/year
samp1 <- samp %>% 
  filter(!is.na(Month) | !Month < 1 | !Month > 12 |
         !is.na(Year) | !Year < 2023 | !Year > 2025)
```
make sure the crime count categories are numeric - make a list of them first
```{r}
# Convert crime category columns to numeric, allowing for negative and positive numbers and 0
crime_categories <- c("Murder", "Rape", "Robbery", 
                  "Agg.Assault", "Burglary", "Larceny", "MVT", "Arson")
samp <- samp %>%
  mutate(across(all_of(crime_categories), ~as.numeric(replace(., . == "NR", NA))))
```
drop all cases by this point that have NA's in their crime Counts
first, segregate them to check them out.
```{r}
# Identify rows with NA or "NA" values
obs_with_na <- apply(samp, 1, function(row) any(is.na(row) | row == "NA"))

# Save these observations to a separate data frame for review
df_na <- df[obs_with_na, ]

# Display the data frame with NA or "NA" values
print("Data frame with NA or 'NA' values:")
head(df_na)

# Remove these observations from the original data frame
df_clean <- df[!obs_with_na, ]

# Display the cleaned data frame
print("Cleaned data frame:")
head(df_clean)
```
Display observations with invalid months
```{r}
print("Observations with Invalid Months:")
print(invalid_months)
```
Display observations with invalid years
```{r}
print("Observations with Invalid Years:")
print(invalid_years)
```
Display observations with "NR" values for crime counts
```{r}
print("Observations with 'NR' in Crime Counts:")
print(nr_observations)
```
Function to calculate lag time - in progress
```{r}

```
before we run the audit, let's standardize the agency name column so that it will join with historical UCR data
```{r}
# List of text pieces to remove
remove_list <- c("PD", "Police Department", "Department of Public Safety", "MIP", "Police Bureau", "Police Division", "\\(Clermont\\)", "PD MIP", "Police Bureau MIP")

# Create a regex pattern from the remove list - look more into this - thank you chatgpt
pattern <- paste(remove_list, collapse = "|")

# Remove specified pieces of text from Agency.Name
samp <- samp %>%
  mutate(City = gsub(pattern, "", Agency.Name, ignore.case = TRUE))

# Trim leading and trailing whitespaces
samp <- samp %>%
  mutate(City = trimws(City))

#create a second City column for viz format
samp$City_viz <- samp$City

# drop city to lowercase to match ucr
samp$City <- tolower(df$Name)

# View the modified data frame
head(samp)
```
run wheeler's outliers tests
```{r}
#we will source it first
source("crimeAnomaly.R")

# This function looks at cumulative year to date stats
# so if prior total 100 YTD (averaged over prior_max years)
# and current is 20 YTD total, this would flagged
# this expects the values in per a single PD
ytd_poisson(samp)

# So this will loop over all PDs and return a total count column
# along with a note, if the note says
# "Burglary_ytd1 | MVT_ytd1 | Larceny_pri8"
# That means this police department was flagged for year to date burglary, year to date MVT
# and prior 8 larceny (but not prior 1 larceny)
# sorts the results, so the agencies with the most flagges are at the top of the dataframe
metrics_allpd(samp)

# This function looks at the average of the prior k months
# so if average of prior 8 is 30, and current month is 10
# this will likely flag
# can set prior_max to 1 to just look at current vs prior month
priork_poisson(samp)
```
reshape data post so we get a clean join to the ucr historical data
```{r}
# Convert Month and Year to a Date column in "%b-%Y" format
samp <- samp %>%
  mutate(Date = as.Date(paste(Year, Month, "01", sep = "-"), "%Y-%b-%d"))

# Format the Date column as "%b-%Y"
samp <- samp %>%
  mutate(Date = format(Date, "%b-%Y"))

# Reshape to long 
samp_long <- samp %>%
  pivot_longer(cols = c("Murder", "Rape", "Robbery", 
                        "Agg Assault", "Burglary", "Larceny", "MVT", "Arson"),
                            names_to = "Crime Type",
                            values_to = "Count")

# Reshape to wide format with Date as columns
samp_wide <- samp_long %>%
  pivot_wider(names_from = Date, values_from = Count)

#let's see what we have now
head(samp_wide)
```

#Now we join the historical data to the sample (technically vice versa)
C:\OneDrive\OneDrive - ahdatalytics.com\Clients\Real Time Crime Index\Historical Data
UCR Historical Data 2018-2022.csv
```{r}
file_path <- "C:\\OneDrive\\OneDrive - ahdatalytics.com\\Clients\\Real Time Crime Index\\Historical Data\\UCR Historical Data 2018-2022.csv"
ucr_data <- read.csv(file_path)
```
reshape the ucr data wide before we join to the sample
```{r}
# Standardize the month first before making the Date column
standardize_month(ucr_data)

# Convert Month and Year to a Date column in "%b-%Y" format
ucr_data <- ucr_data %>%
  mutate(Date = as.Date(paste(Year, Month, "01", sep = "-"), "%Y-%b-%d"))

# Format the Date column as "%b-%Y"
ucr_data <- ucr_data %>%
  mutate(Date = format(Date, "%b-%Y"))

# Reshape to long 
ucr_long <- ucr_data %>%
  pivot_longer(cols = c("Murder", "Rape", "Robbery", 
                        "Agg Assault", "Burglary", "Larceny", "MVT", "Arson"),
                            names_to = "Crime Type",
                            values_to = "Count")

# Reshape to wide format with Date as columns
ucr_wide <- ucr_long %>%
  pivot_wider(names_from = Date, values_from = Count)

#let's see what we have now
head(ucr_wide)
```
now let's join using city and state
```{r}
samp_full_wide <- right_join(ucr_wide, samp_wide, by = c("City", "State"))
```
make sure city name is in Format. 
ILLINOIS and lowercase transformations -app view purposes
```{r}

```

```{r}

```